---
title: new test post
author: Mayan Navon
date: '2020-04-21'
slug: test-post
categories:
  - IAT
  - Israel
  - Age-IAT
tags: []
always_allow_html: true
self.contained: true
---


```{r load-packages, message=FALSE, warning=FALSE, include=FALSE}
options(blogdown.ext = ".Rmd")
# basedir <- 'C:/Users/Lenovo/Google Drive/School/PhD/Lab projects/iat israel/pi-blog-international/images/' 
# outdir <- '/pi-blog-international/images/'
# knitr::opts_knit$set(base.dir = basedir, base.url = outdir, child.path = outdir, url.dir = 'images')

pacman::p_load(ggplot2,rlang,rmdformats,magrittr,dplyr,knitr,kableExtra,tidyr,jmv, ggpubr, table1, ggstatsplot, data.table, effectsize, correlation, see, report) 
#source("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Pilots\\analysis\\RScriptsForAnalysis\\yba.funcs.R")

```

```{r load-data, message=FALSE, warning=FALSE, include=FALSE}

age <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\age\\age_iat_wide.csv") %>% filter(year >= 2009)
ethnic <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\ethnicity\\ethnic_iat_wide.csv") %>% filter(year >= 2009)
# gender_science <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\gender_science\\gender_science_iat_wide.csv") #%>% filter(year >= 2009)
# race <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\race\\race_iat_wide.csv")# %>% filter(year >= 2009)
# sexuality <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\sexuality\\sexuality_iat_wide.csv") #%>% filter(year >= 2009)
# weight <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\weight\\weight_iat_wide.csv") #%>% filter(year >= 2009)
# skin <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\iat israel\\skin\\skin_iat_wide.csv") #%>% filter(year >= 2009)

```

## Individual Task Results: Age 

<mark>Total N = `r nrow(age)`

Participation per year (N):

```{r message=FALSE, warning=FALSE, include=FALSE}

age %>% group_by(year) %>% tally() %>% kable() %>%
  kable_styling(bootstrap_options = c("striped", "condensed", position = "left", full_width = F))
# 
# 
# ggbarplot(cc, x = "year", y = "n",
#   fill = "lightblue", color = "black", 
#   label = TRUE, lab.col = "white", lab.pos = "in")

```

#### **Evaluation measures**

*Note*: Higher values indicate stronger bias in favor of the dominant category (young people).

* *IAT*: There is a strong bias in favor of Young people

```{r echo=FALSE, message=FALSE, warning=FALSE}

pp <- ggdensity(age,
       x = "IAT",
       y = "..density..",
       fill = "lightblue",
       xlab = "IAT", 
       add = "mean",
       size = 1, 
       rug = TRUE
)
pp <- pp + font("title", size = 20, color = "black", face = "bold")+
font("legend.text", size = 16, color = "black", face = "bold")+
font("xlab", size = 18, color = "black", face = "bold")+
font("ylab", size = 18, color = "black", face = "bold")+
font("xy.text", size = 15, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")
pp

```

* *Self-reported preference between the categories*: There is a strong bias in favor of Young people

```{r echo=FALSE, message=FALSE, warning=FALSE}

pp <- ggdensity(age,
       x = "att",
       y = "..density..",
       fill = "lightblue",
       xlab = "Preference", 
       add = "mean",
       size = 1, 
       rug = TRUE
)
pp <- pp + font("title", size = 20, color = "black", face = "bold")+
font("legend.text", size = 16, color = "black", face = "bold")+
font("xlab", size = 18, color = "black", face = "bold")+
font("ylab", size = 18, color = "black", face = "bold")+
font("xy.text", size = 15, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")
pp

```

* *Self-reported feeling thermometer (difference score)*: There is a small bias in favor of Young people. The difference between the current and previous measure might result from the nature of each measure: while the Self-reported preference measure directly compares Young and Old people (as does the IAT), the thermometer does not compare the two categories directly. A comparative context might facilitate bias.

```{r echo=FALSE, message=FALSE, warning=FALSE}

pp <- ggdensity(age,
       x = "diff_thermo",
       y = "..density..",
       fill = "lightblue",
       xlab = "Thermometer difference", 
       add = "mean",
       size = 1, 
       rug = TRUE
)
pp <- pp + font("title", size = 20, color = "black", face = "bold")+
font("legend.text", size = 16, color = "black", face = "bold")+
font("xlab", size = 18, color = "black", face = "bold")+
font("ylab", size = 18, color = "black", face = "bold")+
font("xy.text", size = 15, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")
pp

```

* *Correlation between the measures*: There is a small automatic-deliberate correlation, and a strong relation between the two self-report measures.

```{r echo=FALSE, message=FALSE, warning=FALSE}

age %>% corrMatrix(vars = vars(IAT,att,diff_thermo))

```

* *IAT block-order*

```{r echo=FALSE, message=FALSE, warning=FALSE}

age1 <- age %>% drop_na(blockOrder)

pp <- ggstatsplot::ggbetweenstats(
  data = age1,
  x = blockOrder,
  y = IAT,
  xlab = "Block-order", # label for the x-axis variable
  ylab = "IAT", # label for the y-axis variable  
  messages = FALSE,
  notch = TRUE, # show notched box plot
  mean.plotting = TRUE, # whether mean for each group is to be displayed
  mean.ci = TRUE, # whether to display confidence interval for means
  type = "parametric", # which type of test is to be run
  effsize.type = "biased",
  k = 3, # number of decimal places for statistical results  
) + # further modification outside of ggstatsplot
  ggplot2::coord_cartesian(ylim = c(-2, 2))
pp <- pp + font("title", size = 20, color = "black", face = "bold")+
font("legend.text", size = 16, color = "black", face = "bold")+
font("xlab", size = 18, color = "black", face = "bold")+
font("ylab", size = 18, color = "black", face = "bold")+
font("xy.text", size = 15, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")
pp

```

#### **Evaluation measures by year**

The deliberate and automatic preference between the groups stays the same (and strong) throughout the years.
*Note*: The dashed line marks the point of no preference between the two categories (Old people and Young people).

```{r echo=FALSE, message=FALSE, warning=FALSE}

age1 <- age %>% drop_na(year)
age1$Year <- as.factor(age$year)

p1 <- ggviolin(age1, x = "Year",
      y = c("IAT"),
      combine = TRUE,
      ylab = "IAT", 
      color = "black",
      fill = "lightblue",
      add = c("mean_ci"),
      draw_quantiles = c(0.25,0.5,0.75)
      )
p1 <- p1 + geom_hline(yintercept=0, linetype="dashed", 
                color = "darkblue", size=1.2)

p1 <- p1 + 
font("title", size = 14, color = "black", face = "bold")+
font("legend.text", size = 13, color = "black", face = "bold")+
font("xlab", size = 14, color = "black", face = "bold")+
font("ylab", size = 14, color = "black", face = "bold")+
font("xy.text", size = 10, color = "black")
p1 <- p1 +theme(legend.position="none")

p2 <- ggviolin(age1, x = "Year",
      y = c("diff_thermo"),
      combine = TRUE,
      ylab = "Thermometer", 
      color = "black",
      fill = "lightgreen",
      add = c("mean_ci"),
      draw_quantiles = c(0.25,0.5,0.75)
      )
p2 <- p2 + geom_hline(yintercept=0, linetype="dashed", 
                color = "darkgreen", size=1.2)

p2 <- p2 + 
font("title", size = 14, color = "black", face = "bold")+
font("legend.text", size = 13, color = "black", face = "bold")+
font("xlab", size = 14, color = "black", face = "bold")+
font("ylab", size = 14, color = "black", face = "bold")+
font("xy.text", size = 10, color = "black")
p2 <- p2 +theme(legend.position="none")

p3 <- ggviolin(age1, x = "Year",
      y = c("att"),
      combine = TRUE,
      ylab = "Preference", 
      color = "black",
      fill = "yellow",
      add = c("mean_ci"),
      draw_quantiles = c(0.25,0.5,0.75)
      )
p3 <- p3 + geom_hline(yintercept=4, linetype="dashed", 
                color = "orange", size=1.2)

p3 <- p3 + 
font("title", size = 14, color = "black", face = "bold")+
font("legend.text", size = 13, color = "black", face = "bold")+
font("xlab", size = 14, color = "black", face = "bold")+
font("ylab", size = 14, color = "black", face = "bold")+
font("xy.text", size = 10, color = "black")
p3 <- p3 +theme(legend.position="none")

ggarrange(p1,                                                 
          ggarrange(p2, p3, ncol = 2), 
          nrow = 2)                                       

```

