---
title: 'Biases in Israel'
author: 'Mayan Navon'
date: '2020-10-01'
slug: biases-in-israel
categories:
  - Israel
  - Age-IAT
  - Ethnicity-IAT
  - Gender-Science IAT
  - IAT
  - Nationalism-IAT
  - Race-IAT
  - Sexuality-IAT
  - Skin-IAT
  - Weight-IAT
tags: []
always_allow_html: true
self.contained: true
output:
  blogdown::html_page:
      toc: false
type: ''
subtitle: ''
description: 'Read about the implict and explicit biases in Israel'
image: ''  
geometry: left=1cm, right=1cm, top=1cm, bottom=1cm
classoption: landscape
---

```{r setup, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	comment = NA,
	prompt = FALSE,
	tidy = TRUE
)
library(knitr)

options(max.print="275")
options(width = 1430)
opts_knit$set(width=200) 
knitr::opts_chunk$set(fig.width=25, fig.height=10) 

```

```{r load packages, message=FALSE, warning=FALSE, include=FALSE}

pacman::p_load(data.table,ggplot2,rlang,rmdformats,magrittr,dplyr,knitr,kableExtra,tidyr,jmv, ggpubr)
#source("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Pilots\\analysis\\RScriptsForAnalysis\\yba.funcs.R")

```

```{r load data, include=FALSE}

age <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\age\\age_iat_wide.csv") %>% filter(year >= 2009)
ethnic <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\ethnicity\\ethnic_iat_wide.csv") %>% filter(year >= 2009)
gender_science <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\gender_science\\gender_science_iat_wide.csv") %>% filter(year >= 2009)
race <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\race\\race_iat_wide.csv") %>% filter(year >= 2009)
sexuality <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\sexuality\\sexuality_iat_wide.csv") %>% filter(year >= 2009)
weight <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\weight\\weight_iat_wide.csv") %>% filter(year >= 2009)
skin <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\skin\\skin_iat_wide.csv") %>% filter(year >= 2009)
usa <- fread("C:\\Users\\Lenovo\\Google Drive\\School\\PhD\\Lab projects\\pi international data\\Israel\\usa\\usa_iat_wide.csv") %>% filter(year >= 2009)

```

```{r}

nall <- nrow(race) + nrow(skin) + nrow(ethnic) + nrow(age) + nrow(weight) + nrow(sexuality) + nrow(usa) + nrow(gender_science) 

```



The [Israeli site](https://implicit.harvard.edu/implicit/israel/demo.jsp) offers eight IAT topics for the participants to complete: 

* Race: ‘Black - White’ IAT (N = `r nrow(race)`)
* Ethnicity: ‘Ashkenazi - Mizrahi’ IAT (N = `r nrow(ethnic)`)
* Skin-tone: ‘Light Skin - Dark Skin’ IAT (N = `r nrow(skin)`)
* Age: ‘Young - Old’ IAT (N = `r nrow(age)`)
* Sexuality: ‘Gay - Straight’ IAT (N = `r nrow(sexuality)`)
* Weight: ‘Fat - Thin’ IAT (N = `r nrow(weight)`)
* Nationalism: ‘USA - Own Country’ IAT (N = `r nrow(usa)`) 
* Gender - Science: ‘Male+Science - Female+Liberal arts’ IAT (N = `r nrow(gender_science)`)

Each tasks also included 1-2 self-reported evaluation measures:

* Self-Reported Direct Preference: Preference between the dominant and non-dominant category (scale: 1-7)
* Feeling Thermometer: Cold to warm feelings toward each category (scale: 0-10) 

The tasks also included secondary measures such as perception measures and demographic items. You can check out the data on OSF!


We analyzed the results of each task, from 12 years of data collection (2009-2020) and a total of `r nall` participants. 
We found that *automatic bias* is robust and persistent: The IAT scores from all tasks revealed that Israelis **strongly** prefer  the dominant social groups (e.g., White people, Ashkenazi people, Young people) and not the non-dominant social groups (Black people, Mizrahi people, Old people).


## IAT Scores per Task  

All tasks point to a dominant > non-dominant bias. The strongest automatic bias among Israeli participants is regarding **Nationalism (Own Country > USA)**, **Age (Young > Old)**, and **Gender-Science stereotypes (Male+Science - Female+Liberal arts)**, and the smallest automatic bias is regarding **Sexuality (Straight > Gay people)** and **Weight (Thin > Fat)**:

```{r}

# compute es - iat

aget <- ttestOneS(age, vars = vars(IAT), testValue = 0, effectSize = T)
aget <- as.data.frame(aget$ttest)
aged <- round(aget[[6]],2)

racet <- ttestOneS(race, vars = vars(IAT), testValue = 0, effectSize = T)
racet <- as.data.frame(racet$ttest)
raced <- round(racet[[6]],2)

skint <- ttestOneS(skin, vars = vars(IAT), testValue = 0, effectSize = T)
skint <- as.data.frame(skint$ttest)
skind <- round(skint[[6]],2)

weightt <- ttestOneS(weight, vars = vars(IAT), testValue = 0, effectSize = T)
weightt <- as.data.frame(weightt$ttest)
weightd <- round(weightt[[6]],2)

ethnict <- ttestOneS(ethnic, vars = vars(IAT), testValue = 0, effectSize = T)
ethnict <- as.data.frame(ethnict$ttest)
ethnicd <- round(ethnict[[6]],2)

sexualityt <- ttestOneS(sexuality, vars = vars(IAT), testValue = 0, effectSize = T)
sexualityt <- as.data.frame(sexualityt$ttest)
sexualityd <- round(sexualityt[[6]],2)

usat <- ttestOneS(usa, vars = vars(IAT), testValue = 0, effectSize = T)
usat <- as.data.frame(usat$ttest)
usad <- round(usat[[6]],2)

gen_sci_t <- ttestOneS(gender_science, vars = vars(IAT), testValue = 0, effectSize = T)
gen_sci_t <- as.data.frame(gen_sci_t$ttest)
gen_sci_d <- round(gen_sci_t[[6]],2)

all_ds <- c(aged,raced,skind,weightd,ethnicd,sexualityd,usad,gen_sci_d) %>% as.data.frame()
colnames(all_ds) <- "es"

all_ds$Task <- c("Age","Race","Skin","Weight","Ethnicity","Sexuality","Nationalism","Gender-Science")

```

```{r}

pp <- ggdotchart(all_ds, x = "Task", y = "es",
           color = "Task",                                # Color by groups
           ylab = "Cohen's d",
           #palette = c("#00AFBB", "#E7B800", "#FC4E07"), # Custom color palette
           sorting = "descending",                       # Sort value in descending order
           add = "segments",                             # Add segments from y = 0 to dots
           rotate = TRUE,                                # Rotate vertically
           #group = "cyl",                                # Order by groups
           dot.size = 30,                                 # Large dot size
           label = round(all_ds$es,3),                        # Add mpg values as dot labels
           font.label = list(color = "white", size = 25, 
                             vjust = 0.5),               # Adjust label parameters
           ggtheme = theme_pubr()                        # ggplot2 theme
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2)

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp

```


## Self-Report Scores per Task

Israelis were often biased *explicitly*, and not just automatically... but, as might be expected, this bias was weaker in its magnitude, compared with the automatic bias. 

All tasks point to a dominant > non-dominant bias. The strongest self-reported bias among Israeli participants is regarding **Gender-Science stereotypes (Male+Science - Female+Liberal arts)** and **Weight (Thin > Fat)** and the smallest automatic bias is regarding **Sexuality (Straight > Gay people)**:

### Self-Reported Direct Preference

*Note*: Not all tasks included the direct self-reported preference measure.

```{r}

# compute es - self reported direct preference

aget <- ttestOneS(age, vars = vars(att), testValue = 4, effectSize = T)
aget <- as.data.frame(aget$ttest)
aged <- round(aget[[6]],2)

racet <- ttestOneS(race, vars = vars(att), testValue = 4, effectSize = T)
racet <- as.data.frame(racet$ttest)
raced <- round(racet[[6]],2)

weightt <- ttestOneS(weight, vars = vars(att), testValue = 4, effectSize = T)
weightt <- as.data.frame(weightt$ttest)
weightd <- round(weightt[[6]],2)

ethnict <- ttestOneS(ethnic, vars = vars(att), testValue = 4, effectSize = T)
ethnict <- as.data.frame(ethnict$ttest)
ethnicd <- round(ethnict[[6]],2)

sexualityt <- ttestOneS(sexuality, vars = vars(att), testValue = 4, effectSize = T)
sexualityt <- as.data.frame(sexualityt$ttest)
sexualityd <- round(sexualityt[[6]],2)

usat <- ttestOneS(usa, vars = vars(att), testValue = 4, effectSize = T)
usat <- as.data.frame(usat$ttest)
usad <- round(usat[[6]],2)

all_ds <- c(aged,raced,weightd,ethnicd,sexualityd,usad) %>% as.data.frame()
colnames(all_ds) <- "es"

all_ds$Task <- c("Age","Race","Weight","Ethnicity","Sexuality","Nationalism")

```

```{r}

pp <- ggdotchart(all_ds, x = "Task", y = "es",
           color = "Task",                                # Color by groups
           ylab = "Cohen's d",
           #palette = c("#00AFBB", "#E7B800", "#FC4E07"), # Custom color palette
           sorting = "descending",                       # Sort value in descending order
           add = "segments",                             # Add segments from y = 0 to dots
           rotate = TRUE,                                # Rotate vertically
           #group = "cyl",                                # Order by groups
           dot.size = 30,                                 # Large dot size
           label = round(all_ds$es,3),                        # Add mpg values as dot labels
           font.label = list(color = "white", size = 25, 
                             vjust = 0.5),               # Adjust label parameters
           ggtheme = theme_pubr()                        # ggplot2 theme
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2)

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp

```

### Feeling thermometer (Difference score)

*Note*: Not all tasks included the feeling thermometer measure.

```{r}

# compute es - thermometer

aget <- ttestOneS(age, vars = vars(diff_thermo), testValue = 0, effectSize = T)
aget <- as.data.frame(aget$ttest)
aged <- round(aget[[6]],2)

racet <- ttestOneS(race, vars = vars(diff_thermo), testValue = 0, effectSize = T)
racet <- as.data.frame(racet$ttest)
raced <- round(racet[[6]],2)

skint <- ttestOneS(skin, vars = vars(diff_thermo), testValue = 0, effectSize = T)
skint <- as.data.frame(skint$ttest)
skind <- round(skint[[6]],2)

weightt <- ttestOneS(weight, vars = vars(diff_thermo), testValue = 0, effectSize = T)
weightt <- as.data.frame(weightt$ttest)
weightd <- round(weightt[[6]],2)

usat <- ttestOneS(usa, vars = vars(diff_thermo), testValue = 0, effectSize = T)
usat <- as.data.frame(usat$ttest)
usad <- round(usat[[6]],2)

all_ds <- c(aged,raced,skind,weightd,usad) %>% as.data.frame()
colnames(all_ds) <- "es"

all_ds$Task <- c("Age","Race","Skin","Weight","Nationalism")

```

```{r}

pp <- ggdotchart(all_ds, x = "Task", y = "es",
           color = "Task",                                # Color by groups
           ylab = "Cohen's d",
           #palette = c("#00AFBB", "#E7B800", "#FC4E07"), # Custom color palette
           sorting = "descending",                       # Sort value in descending order
           add = "segments",                             # Add segments from y = 0 to dots
           rotate = TRUE,                                # Rotate vertically
           #group = "cyl",                                # Order by groups
           dot.size = 30,                                 # Large dot size
           label = round(all_ds$es,3),                        # Add mpg values as dot labels
           font.label = list(color = "white", size = 25, 
                             vjust = 0.5),               # Adjust label parameters
           ggtheme = theme_pubr()                        # ggplot2 theme
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2)

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp

```

**Do people get better?**
- We next looked at the scores per year, to see if bias decreases throughout the years.

## Yearly patterns

What we found, is that bias, both automatic and self-reported stays the same throughout the 12 years. Across all years, we can see that self-reported and automatic judgments favor the dominant groups, and that automatic bias is much larger.

```{r}

age <- age %>% drop_na(year)
age$Year <- as.factor(age$year)

race <- race %>% drop_na(year)
race$Year <- as.factor(race$year)

skin <- skin %>% drop_na(year)
skin$Year <- as.factor(skin$year)

ethnic <- ethnic %>% drop_na(year)
ethnic$Year <- as.factor(ethnic$year)

weight <- weight %>% drop_na(year)
weight$Year <- as.factor(weight$year)

sexuality <- sexuality %>% drop_na(year)
sexuality$Year <- as.factor(sexuality$year)

usa <- usa %>% drop_na(year)
usa$Year <- as.factor(usa$year)

gender_science<- gender_science %>% drop_na(year)
gender_science$Year <- as.factor(gender_science$year)

```

```{r}

race_iat <- race %>% select(session_id,IAT,Year) %>% mutate(Task = "Race")
skin_iat <- skin %>% select(session_id,IAT,Year) %>% mutate(Task = "Skin")
age_iat <- age %>% select(session_id,IAT,Year) %>% mutate(Task = "Age")
ethnic_iat <- ethnic %>% select(session_id,IAT,Year) %>% mutate(Task = "Ethnicity")
sex_iat <- sexuality %>% select(session_id,IAT,Year) %>% mutate(Task = "Sexuality")
usa_iat <- usa %>% select(session_id,IAT,Year) %>% mutate(Task = "Nationalism")
gender_science_iat <- gender_science %>% select(session_id,IAT,Year) %>% mutate(Task = "Gender-Science")
weight_iat <- weight %>% select(session_id,IAT,Year) %>% mutate(Task = "Weight")

all_iat <- bind_rows(race_iat,skin_iat,ethnic_iat,age_iat,sex_iat,usa_iat,gender_science_iat,weight_iat)


race_att <- race %>% select(session_id,att,Year) %>% mutate(Task = "Race")
#skin_att <- skin %>% select(session_id,att,Year) %>% mutate(Task = "Skin")
age_att <- age %>% select(session_id,att,Year) %>% mutate(Task = "Age")
ethnic_att <- ethnic %>% select(session_id,att,Year) %>% mutate(Task = "Ethnicity")
sex_att <- sexuality %>% select(session_id,att,Year) %>% mutate(Task = "Sexuality")
usa_att <- usa %>% select(session_id,att,Year) %>% mutate(Task = "Nationalism")
#gender_science_att <- gender_science %>% select(session_id,att,Year) %>% mutate(Task = "Gender-Science")
weight_att <- weight %>% select(session_id,att,Year) %>% mutate(Task = "Weight")

all_att <- bind_rows(race_att,ethnic_att,age_att,sex_att,usa_att,weight_att)
all_att$att_rs <- all_att$att - 4

race_diff_thermo <- race %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Race")
skin_diff_thermo <- skin %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Skin")
age_diff_thermo <- age %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Age")
#ethnic_diff_thermo <- ethnic %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Ethnicity")
#sex_diff_thermo <- sexuality %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Sexuality")
usa_diff_thermo <- usa %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Nationalism")
#gender_science_diff_thermo <- gender_science %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Gender-Science")
weight_diff_thermo <- weight %>% select(session_id,diff_thermo,Year) %>% mutate(Task = "Weight")

all_diff_thermo <- bind_rows(race_diff_thermo,skin_diff_thermo,age_diff_thermo,usa_diff_thermo,weight_diff_thermo)

```

```{r get es per task & year}

library(magrittr)
library(effectsize)

# iat

get_es_iat <- function(x)
{
  es <- x$IAT %>% cohens_d()
  es <- round((es),2)
  es_list <- list(es = es)
}

df <- all_iat %>% drop_na() %>% group_by(Task,Year) %>% nest() %>% drop_na()

iat_df <- df %>% mutate(params = purrr::map(.x = data, .f = ~get_es_iat(.x))) %>% 
        unnest_wider(params) %>% select(-data) %>% unnest_wider(es)

# self-reported preference

get_es_sr <- function(x)
{
  es <- x$att_rs %>% cohens_d()
  es <- round((es),2)
  es_list <- list(es = es)
}

df <- all_att %>% drop_na() %>% group_by(Task,Year) %>% nest() %>% drop_na()

att_df <- df %>% mutate(params = purrr::map(.x = data, .f = ~get_es_sr(.x))) %>% 
        unnest_wider(params) %>% select(-data) %>% unnest_wider(es)

# self-reported thermometer

get_es_thr <- function(x)
{
  es <- x$diff_thermo %>% cohens_d()
  es <- round((es),2)
  es_list <- list(es = es)
}

df <- all_diff_thermo %>% drop_na() %>% group_by(Task,Year) %>% nest() %>% drop_na()

thermo_df <- df %>% mutate(params = purrr::map(.x = data, .f = ~get_es_thr(.x))) %>% 
        unnest_wider(params) %>% select(-data) %>% unnest_wider(es)

```


*Note*: The dashed line marks the point of no preference between the two groups.

### IAT

```{r}

iat_df$Year = factor(iat_df$Year, c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020"))

pp <- ggdotchart(iat_df, x = "Year", y = "Cohens_d",
           color = "Task",
           ylab = "Cohen's d",
           sorting = "none",
           order = NULL,
           combine = F,
           add = "segments",
           facet = "Task",
           rotate = F,
           dot.size = 30,
           label = round(iat_df$Cohens_d,3),
           font.label = list(color = "white", size = 25,
                             vjust = 0.5),
           ggtheme = theme_pubr()
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2) + theme(axis.text.x=element_text(angle = 22))

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp


```

### Self-Reported Direct Preference

```{r}

att_df$Year = factor(att_df$Year, c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020"))

pp <- ggdotchart(att_df, x = "Year", y = "Cohens_d",
           color = "Task",
           ylab = "Cohen's d",
           sorting = "none",
           order = NULL,
           combine = F,
           add = "segments",
           facet = "Task",
           rotate = F,
           dot.size = 30,
           label = round(att_df$Cohens_d,3),
           font.label = list(color = "white", size = 25,
                             vjust = 0.5),
           ggtheme = theme_pubr()
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2) + theme(axis.text.x=element_text(angle = 22))

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp


```

### Self-Report: Feeling Thermometer Difference

```{r}

thermo_df$Year = factor(thermo_df$Year, c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020"))

pp <- ggdotchart(thermo_df, x = "Year", y = "Cohens_d",
           color = "Task",
           ylab = "Cohen's d",
           sorting = "none",
           order = NULL,
           combine = F,
           add = "segments",
           facet = "Task",
           rotate = F,
           dot.size = 30,
           label = round(thermo_df$Cohens_d,3),
           font.label = list(color = "white", size = 25,
                             vjust = 0.5),
           ggtheme = theme_pubr()
           )

pp <- pp + geom_hline(yintercept = 0, linetype = "dashed", color = "darkblue", 
    size = 1.2) + theme(axis.text.x=element_text(angle = 22))

pp <- pp + font("title", size = 30, color = "black", face = "bold")+
font("legend.text", size = 26, color = "black", face = "bold")+
font("xlab", size = 28, color = "black", face = "bold")+
font("ylab", size = 28, color = "black", face = "bold")+
font("xy.text", size = 25, color = "black", face = "bold")
pp <- pp +theme(legend.position="none")

pp

```


## The Ethnicity IAT

This IAT uses categories unique to Israel. Let's inspect it further. 
Do people from different ethnicity have different levels of automatic ethnicity-related bias?

We'll start by examining the participation per ethnicity group:

```{r}

yy <- ethnic %>% drop_na (ethnic.c) %>% group_by(ethnic.c) %>% tally()
yyt <- as.data.frame(t(yy$n))
colnames(yyt) <- c("5+ Zabar", "Arab countries", "East Asia", "East Europe", "Ethiopia", "Mixed (other)", 
    "Mixed Miz/Ash", "North Africa", "North America", "Other", "Russia", "South American", "West Europe")

yyt %>% kable() %>% kable_styling(bootstrap_options = c("striped", 
    "hover", "condensed", full_width = F, position = "left"))

```

Now let's look at the IAT scores per ethnicity group:

```{r}

ethnic1 <- ethnic %>% drop_na(ethnic.c)

pp <- ggviolin(ethnic1, x = "ethnic.c",
      y = c("IAT"),
      combine = TRUE,
      ylab = "IAT", 
      xlab = "Ethnicity",
      color = "black",
      fill = "lightblue",
      add = c("mean_ci"),
      draw_quantiles = c(0.25,0.5,0.75)
      )
pp <- pp + geom_hline(yintercept=0, linetype="dashed", 
                color = "darkblue", size=1.2)

pp <- pp + 
font("title", size = 24, color = "black", face = "bold")+
font("legend.text", size = 23, color = "black", face = "bold")+
font("xlab", size = 24, color = "black", face = "bold")+
font("ylab", size = 24, color = "black", face = "bold")+
font("xy.text", size = 20, color = "black")
pp <- pp +theme(legend.position="none")
pp


```

We can see that, indeed, the automatic bias varies across the different ethnicity. Still, *all ethnicity groups* in the Israeli society show the same direction of bias.



## Summary

* Across all tasks, Israelis were biased, explicitly and automatically, in favor of the dominant social groups.
* Across all tasks, the bias (preference) was much larger for the automatic than the self-reported (explicit) judgment.
* The bias stayed the same across the 12 years of data collection.
* IAT scores: The strongest bias among Israeli participants was regarding Age (Young > Old people) and the smallest bias was regarding Sexuality (Straight > Gay people).

## Show me more!

All the data, as well as the full analysis report, are available on OSF **INSERT LINK HERE**.
